Ages of Planets and Stars Observed by Kepler

### 2024-04-02 paper/ms.tex
update ms

Using our adopted effective temperatures and rotation periods, we ran 
\texttt{gyro-interp} over an age grid of 0 to 5\,Gyr, linearly spaced
over 500 grid points.
\texttt{gyro-interp} itself is only calibrated out to 4\,Gyr, the
adopted age of the M67 cluster \citep[see][]{Dungee_2022,Gruner_2023}.
We used the spin-down law between the reference open clusters by
interpolating using piecewise cubic hermite interpolating polynomials,
and integrated Equation~1 of \citep{Bouma_2023} at the default grid
resolution.

\subsection{Star Filters}
We calculated gyrochrone age posteriors for all \nuniqstarsantosrot\
stars with reported rotation periods and effective temperatures.  To
select the subset stars for which we expect these age to in fact be
valid, we then build a set of filters and quality flags, and apply
them in the sub-analyses that follow.  We defined these filters as
follows, each of which is included in Tables~\ref{tab:stars}
and~\ref{tab:planets}.

@@

The rotation periods for the stellar and planetary samples are shown
in Figure~\ref{fig:prot_vs_teff}.
Out of the \nkeplerstars\ stars observed by Kepler, requiring reported
rotation periods yielded \nuniqstarsantosrotgyroappl.  Further
requiring these rotating stars to be apparently single, near the main
sequence, with $\log g$$>$4.2, RUWE$<$1.4, 3800$<$$T_{\rm
eff}$/K$<$6200, and not flagged as eclipsing binaries yielded
\nuniqstarsantosrotgyroappl\ rotators amenable for gyrochronology.

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{prot_teff_Santos19_Santos21_clean0.pdf}
		}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{koi_mean_prot_teff_koi_X_S19S21dquality_keepgrazing.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Rotation periods for Kepler stars (left) and planets (right)}.
    The periods are from \citet{Santos_2019} and \citet{Santos_2021}
    (\citetalias{Santos_2019} and \citetalias{Santos_2021}), and the
    temperatures are primarily from \citet{Berger_2020a_catalog}.  The
    gray lines are ``mean fits'' from \citet{Bouma_2023} to the slow
    rotation sequences of open clusters.
    The stellar sample shows only apparently single stars near the
    main sequence with $\log g$$>$4.2, RUWE$<$1.4, 3800$<$$T_{\rm
    eff}$/K$<$6200, that are not flagged as eclipsing binaries.  The
    planet sample has the same stellar cuts, and shows confirmed and
    candidate planets from the cumulative KOI table, requiring them to
    have a MES$>$10.
    % koi_mean_prot_teff_koi_X_S19S21dquality_keepgrazing.png
    % minus
    % koi_mean_prot_teff_koi_X_S19S21dquality_dropgrazing.png
    %Grazing objects ($b$>0.8) include 213 of the planet candidates
    %displayed, and 136 of the host stars.
	}
	\label{fig:prot_vs_teff}
\end{figure*}

\subsection{Ages for stars in the Kepler field}
%TODO: show parameter distributions?

>>>>

Using our adopted effective temperatures and rotation periods, we ran
\texttt{gyro-interp} over an age grid of 0 to 4\,Gyr, linearly spaced
over 500 grid points.  \texttt{gyro-interp} itself is only calibrated
out to 4\,Gyr, the adopted age of the M67 cluster
\citep[see][]{Dungee_2022,Gruner_2023}.  To estimate the stellar
spin-down rate as a function of time and stellar temperature, this
framework interpolates between reference open clusters using smooth
cubic hermite polynomials.  We calculate the probability of the
rotation-based age $t_{\rm gyro}$, given the observed rotation
periods, effective temperatures, and their uncertainties by
integrating Equation~1 of \citep{Bouma_2023} at the default grid
resolution.  This procedure marginalizes over the intrinsic age and
temperature dependent scatter that is observed in the cluster
sequences to yield statistical age uncertainties.

\subsection{Star Filters}
\label{subsec:filters}
We calculated gyrochrone age posteriors for all \nuniqstarsantosrot\
stars with reported rotation periods and effective temperatures.  To
select the subset stars for which we expect these age to in fact be
valid, we then build a set of filters and quality flags, and apply
them in the sub-analyses that follow.  We defined these filters as
follows, each of which is included in Tables~\ref{tab:stars}
and~\ref{tab:planets}.

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{prot_teff_Santos19_Santos21_clean0.pdf}
		}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{koi_mean_prot_teff_koi_X_S19S21dquality_keepgrazing.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Rotation periods for Kepler stars (left) and planets (right)}.
    The periods are from \citet{Santos_2019} and \citet{Santos_2021}
    (\citetalias{Santos_2019} and \citetalias{Santos_2021}), and the
    temperatures are primarily from \citet{Berger_2020a_catalog}.  The
    gray lines are ``mean fits'' from \citet{Bouma_2023} to the slow
    rotation sequences of open clusters.  The stellar sample shows
    only apparently single stars near the main sequence with $\log
    g$$>$4.2, RUWE$<$1.4, 3800$<$$T_{\rm eff}$/K$<$6200, that are not
    flagged as eclipsing binaries.  The planet sample has the same
    stellar cuts, and shows confirmed and candidate planets from the
    cumulative KOI table, requiring them to have a MES$>$10.
	}
	\label{fig:prot_vs_teff}
\end{figure*}

The rotation periods for the stellar and planetary samples are shown
in Figure~\ref{fig:prot_vs_teff}.  Out of the \nkeplerstars\ stars
observed by Kepler, requiring reported rotation periods yielded
\nuniqstarsantosrotgyroappl.  Further requiring these rotating stars
to have temperatures 3800$<$$T_{\rm eff}$/K$<$6200, which is the range
over which \texttt{gyro-interp} returns finite age posteriors,
yields \nuniqstarsantosrotteffcut\ stars.
Further requiring them to be apparently single, near the main
sequence, with $\log g$$>$4.2, RUWE$<$1.4, and not flagged as
eclipsing binaries yielded \nuniqstarsantosrotgyroappl\ rotators
amenable for gyrochronology.

@@

\subsection{Age distribution}

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{hist_samples_koi_gyro_ages_hist_field_gyro_ages_20230529_maxage3200.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
		{\bf Age distribution of stars and planet-hosts in the Kepler field
    inferred from rotation periods}.
    A histogram of samples from the age posteriors for all Kepler
    stars with reported rotation periods and temperatures is shown in
    light gray;
    the age calculation was actually valid for stars that meet the criteria
    discussed in Section~\ref{subsec:filters} (dark blue).
		\label{fig:hist_tgyro}
	}
\end{figure*}

Figure~\ref{fig:hist_tgyro} shows the gyrochone age distributions
of stars and planet-hosts in the Kepler field.
We truncated the plots near $3$\,Gyr due to arguments that suggest the
detectability of rotation signals is near-unity up to this age for Sun-like
stars, and decreases for older ages \citep{2022ApJ...937...94M}.
Both the stellar and planetary samples point toward a star
formation rate in the Kepler field that has slowed over the past
few gigayears.
Imagine we were to group stars between 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr, and 
label them ``young'', ``intermediate-age'', and ``old''.
A simple counting exercise from these distributions tells us that there are \ratioobtoybstars\ times as many 
old stars in the Kepler field as young stars.
Similarly, there are \ratioobtoybplanets\ times as many old planet hosts as
young planet hosts.

{\it LGB pending words: touch on findings on MW SFH from chemistry,
isochrones, and asteroseismology}.
